---
title: "R Notebook"
output: html_notebook
---
Here you'll generate your own data to make sure you understand what PCA is doing

Generate 4 variables W, X, Y, and Z

X and Y should not be correlated

They are independent
```{r}
library(factoextra)
library(tidyverse)
```
```{r}
my_table = data.frame(w, x, y, z)
my_table
```
```{r}
?data.frame
```

W and X should have a mild correlation ( < 0.5)
Y and Z should have a mild correlation ( > 0.9)
```{r}
set.seed(42)
x = rnorm(10, 2, 1)
noise = rnorm(10, 0, 2)
```
```{r}

```


```{r}
w = 2*x + noise
cor(x,w)
```


```{r}
y = rnorm(10, 6, 1)
error = rnorm(10, 6, 0.1)

z = y/2 + error
cor(y, z)
```

Here, it's important to note that: The correction of two variables A and B is not determined by the linear regression ğµ=0.4Ã—ğ‘‹ does not mean that the correlation cor(A,B) = 0.4 Correction is factor of the noise in the linear regressio. For example, for ğµ=0.4Ã—ğ‘‹+ğœ– , the larger the noise component, the samller will be the correction between A and B


Generate a variable outcome as a linear combination of W, X, Y, and Z
i.e., choose values for the coefficients  ğ›½0 ,  ğ›½1 , etc.. and compute outcome
ğ‘œğ‘¢ğ‘¡ğ‘ğ‘œğ‘šğ‘’=ğ›½0+ğ›½1Ã—ğ‘Š+ğ›½2Ã—ğ‘‹+ğ›½3Ã—ğ‘Œ+ğ›½4Ã—ğ‘

```{r}
model_1 = lm(x~ y + w:z, data = my_table)
summary(model_1)
```

```{r}
outcome = 1 + 3 * w + 4 * x + 5 * y + 6 * z
outcome
```
```{r}
model_4 = lm(outcome~., data = my_table)
summary(model_4)
```


Model your outcome using W, X, Y, and Z.
Do your results match your model params?

```{r}
model_2 = lm(w~z + y:z, data = my_table)
summary(model_2)
```
```{r}
ggplot(my_table, aes(y, z)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE)
```
```{r}
dif_error = rnorm(10, 0, 1)
output = 4*w + 3*x + 2*y + dif_error
attempt = lm(output ~ x + y + w + z)
summary(attempt)
```

Use PCA to reduce the dimensionality of your dataset
Can you explain why you don't need to include the outcome?

Use the bi-plot to visualize the contributions of your initial variables

```{r}
res_pca = prcomp(my_table, scale=TRUE)
summary(res_pca)

```

```{r}
str(res_pca)
```

```{r}
res_pca$rotation
```

```{r}
res_pca$sdev^2
```

```{r}
res_pca$sdev^2/sum(res_pca$sdev^2)
```


```{r}
res_pca$rotation[4,2]
```


```{r}
fviz_pca_biplot(res_pca)
```
How efficient is the new lower-dimensional space representation at predicting the outcome?
Do your results match your model params?

This new representation is much better at predicting the outcome. Y & Z have a strong correlation greater than 0.95 and are closer together. Meanwhile x and w are barely correlated with a correlation <0.2. This is mirrored in the model above. I think more is gained from seeing this model than the linear regression model. 
